---
title: "Reproducible Research: Peer Assessment 1"
output: 
        html_document:
                keep_md: true
---

knit2html("PA1_template.Rmd")
browseURL("PA1_template.html")

```{r setoptions, echo=FALSE}
knitr::opts_chunk$set(comment = NA)
knitr::opts_chunk$set(error = TRUE)
knitr::opts_chunk$set(fig.align="left")
```

## Loading and preprocessing the data

```{r include=FALSE}
library(knitr)
library(readr)
library(data.table)
library(ggplot2)
library(stringr)
library(mice)

setwd("/Users/donttouch/DScience/ReproducibleResearch/W2")
```

###1. Load the data (i.e.read.csv())

```{r}
if(!file.exists('activity.csv')){
        unzip('activity.zip')
}
activity_dt <- read.csv('activity.csv')
```

###2. Process/transform the data (if neccessary) into a format duitable for your analysis


## What is mean total number of steps taken per day ?

###1. Calculate the total number of steps taken per day 
```{r message=FALSE}
DT<-data.table(activity_dt)
StepsByDay<-DT[,sum(na.omit(steps)),by=.(date)]
setnames(StepsByDay,c("Date","TotalStepsPerDay"))
```
###2. Make a histogram of total number of steps taken each day.
```{r}
StepsByDay[,qplot(TotalStepsPerDay,
        xlab='Steps Per Day',
        ylab='Frequency',
        geom="histogram",
        binwidth=500,
        alpha=I(.2),
        fill=I("blue"),
        col=I("red")
        )]
```

###3. Calculate and report the mean and median of the total number of steps taken per day.

```{r}
out<-StepsByDay[,.(mean(na.omit(TotalStepsPerDay)),median(na.omit(TotalStepsPerDay)))]
```

Mean: `r out$V1`

Median: `r out$V2`


## What is the average daily activity pattern ?

### 1. Make a time series plot(i.e. type = "1") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r fig.width=10}
avgStepsByInterval<-DT[,mean(na.omit(steps)),by=.(interval)]
setnames(avgStepsByInterval,c("interval","avgStepsTakenPerInterval"))
ggplot(avgStepsByInterval, 
        aes(x = interval, y = avgStepsTakenPerInterval)) +
        geom_line() + 
        xlab("Interval(5min)") +
        ylab("Average number of steps taken") +
        theme_bw(base_family="Times")

```

### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
MaxSteps<-which.max(avgStepsByInterval$avgStepsTakenPerInterval)
c<-avgStepsByInterval[MaxSteps,interval]
IntervalWithMaxSteps<-paste(str_sub(c,0,nchar(c)-2),':',str_sub(c,nchar(c)-1,nchar(c)),sep="")

```

Time With Max Steps Taken: `r IntervalWithMaxSteps`

## Inputing Missing Values

### 1. Calculate and Report number of missing values in the dataset (i.e. the total number of Rows with NAs)

```{r}
RowsWithNAs<-md.pattern(activity_dt)
```
```{r echo=FALSE}
RowsWithNAs
NAsFALSE<-rownames(RowsWithNAs)[1]
NAsTRUE<-rownames(RowsWithNAs)[2]
```
The output tells us that `r NAsFALSE` cases are complete and `r NAsTRUE` samples miss Steps measurement.

### 2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc

We can impute missing data by using mice() with mean matching as imputation method. The mice() function would takes care
of the imputing process.

### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r results="hide"}
tempDT <- mice(activity_dt, m=5, maxit=5, meth='mean',seed=500)
```

```{r}
completeDT <- complete(tempDT,1)
```

Check distribution of orginal and imputed data using xyplot plot.

*Shape of the magenta points (imputed) matches the shape of the blue ones (observed).*

```{r}
library(lattice)
xyplot(tempDT,steps ~ interval,pch=18,cex=1)
```

### 4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

*Histogram of the total number of steps taken each day*
*There is no a significant change, except number of step at least range *
```{r message=FALSE }
DTImputed<-data.table(completeDT)
StepsByDayImputed<-DTImputed[,sum(na.omit(steps)),by=(as.Date(date,,'%Y-%m-%d'))]
setnames(StepsByDayImputed,c("Date","TotalStepsPerDay"))
```
```{r echo=FALSE,fig.width=10}
StepsByDayImputed[,qplot(TotalStepsPerDay,
        xlab='Steps Per Day',
        ylab='Frequency',
        geom="histogram",
        main='Imputed Data - Total Number of Steps Taken each Day',
        binwidth=500,
        alpha=I(.2),
        fill=I("blue"),
        col=I("red")
        )]
```

*The mean and median total number of steps per day*

```{r}
out<-StepsByDayImputed[,.(mean(na.omit(TotalStepsPerDay)),median(na.omit(TotalStepsPerDay)))]
```

Mean-Imputed: `r out$V1`

Median-Imputed: `r out$V2`

## Are there differences in activity patterns between weekdays and weekends?

## 1.Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
DTImputed[,dayOfWeek:=ifelse(wday(as.Date(date,'%Y-%m-%d')) %in% c(0,6), 'weekend', 'weekday')]
```

## 2. Make a panel plot

```{r}
avgStepsByDayIntervalImputed<-DTImputed[,AvgSteps:=mean(steps),by=list(interval,dayOfWeek)]
```

```{r fig.width=10}
ggplot(avgStepsByDayIntervalImputed,aes(interval,AvgSteps)) +
geom_line() +
facet_grid(dayOfWeek ~ .) +
xlab("5-min Interval") +
ylab("Avg Number of Steps")
```